Supplementary figures\nCapturing regeneration in models of forest dynamics

### 2023-01-24 manuscript/supplementaryRW.Rmd
updates report from Landis in supplementary

Landscape scale forest simulation model LANDIS-II (Scheller et al. 2007)
with PnET-Succession extension (version 4.1). PnET-Succession is based
on the Biomass Succession extension of Scheller and Mladenoff (2004),
embedding elements of the PnET-II ecophysiology model of Aber et al
(1995) to mechanistically simulate growth as a competition for available
light and water (De Bruijn et al. 2014). PnET-Succession simulates the
competition among cohorts for water and light at a monthly time-step as
a function of photosynthetic processes and maintenance respiration that
are explicitly linked to environmental drivers such as temperature,
precipitation, photosynthetic active radiation (PAR), and CO2
concentration. Regeneration depends on distance from a seed source, soil
water, and subcanopy light, while competition is modeled by partitioning
incoming light through multiple canopy layers. Growth of specific cohort
biomass components is allocated in the following order: non structural
carbon, first, then foliage then root and stem. Biomass growth decreases
as cohorts approach their longevity age but mortality can occur at any
time when carbon reserves production is insufficient to support growth
due to shading, water competition, and/or drought (Gustafson et al.
2015), or by any disturbance or management extension used. For output,
we selected Biomass Community Output Extension version 2.0.1, which
generates output in the format: species age biomass (g / m2) at
specified time steps.

@@

Climate input was generated by randomly selecting entire years from the
monthly data from the CHELSA dataset and collating them one after the
other to create a 2000 years series. The sampling was repeated 10 times
with different random order of years to create the 10 climate
replicates. Selecting a year at a time eliminates inconsistent seasons
(especially summer). We used monthly mean, minimum and maximum
temperatures and precipitation as input for PnET succession extension.
PnET-Succession also requires mean monthly values of PAR during daylight
hours, which we calculated by dividing the monthly rsds data by the
number of daylight seconds in each month for latitude 47 to get W/m2. We
then multiplied this by 4.6 to get μmol/m2/s and multiplied by 0.5 to
get PAR part of the spectrum, based on the approximation by Tsubo et al.
(2005). Leap years were not considered. The PAR distribution across all
sites and years is in range with values that are typical for temperate
regions (Fig. 1).

Figure 1 Histogram of monthly PAR values for all sites and all available
time

CO2 concentration with yearly timestep was added from the concentrations
table available from EEA (2019). For years 0-1750 CO2 concentration was
maintained fixed to 278 ppm. As for the period 1750-1975 there was only
one value each 5 years, we used a constant value for each 5-year period
(Fig. 2). CO2 concentration was added to the climate input file after
randomization of temperature, precipitation and PAR (see above), so it
reflects realistic values for the simulated years. Simulations were run
from year 0 to 1990 but sampled outputs did not exceed year 1910 (see
Simulation length and descriptions further below). Thus, the drastic
increase in CO2 concentration during the second half of the 1900s did
not affect the delivered results.

Figure 2 CO2 concentration values added to the climate data

@@

@@

@@

As LANDIS-II cannot start from bareground, initial communities were set
to include all species defined in the protocol at all sites starting
with 10-years old cohorts of the following species: fasy (Fagus
sylvatica), pabi (Picea abies), abal (Abies alba), cabe (Carpinus
betulus), tico (Tilia cordata), acps (Acer pseudoplatanus), betu (Betula
spp.), frex (Fraxinus excelsior), quer (Quercus spp.), algl (Alnus
glutinosa), pisy (Pinus sylvestris). The species-specific parameter
denoting sexual maturity was set to 10 years equally for all species, so
that seed source of all species is available from the onset of the
simulations at all sites and to maintainseed in the ladscape pool.
Vegetative regeneration (resprouting) was prevented, since this would
have an effect on allometry and regeneration. Some species may be under
or overrepresented because their occurrence in LANDIS-II simulations is
typically due to small-scale disturbances or management which were not
included in the simulations. We have started with a set of parameters we
have used for Petter et al. (2020) which were further refined and
modified for PnET v 4.1. Describe species parameters here...

@@

Simulation area was defined as a rectangle of 10×20 dimension with sites
(i.e., cells) of 1-ha. Each site is an individual ecoregion (i.e., one
cell - as in our case - or group of cells sharing similar climate and
soil conditions, as a way in LANDIS-II to define differences in
biophysical conditions across landscapes) with specific values for soil
quality, topography and climate, based on the data provided. Soil type
was set to LOAM at all sites. Rooting depth was calculated from soil
quality codes (1-5, original values ranged from 1.13 to 4.87)
multiplicated by 205. This produces a soil depth in the range 231 - 998
mm, that are within the typical range for PnET-Succession (maximum 1000
mm). A clear trend emerged - deeper soils resulted in higher biomass,
although with some exceptions (Fig. 3).

Figure 3 Biomass of all species on each site ordered by soil quality
(data from replicate 1). Notice a gap after first longevity period, an
artifact of spin-up from "bare ground"

The ecoregion-specific parameter Precipitation Loss Fraction denotes the
water lost to runoff induced by topographic slope. We estimated this
parameter for each site from topo_data_200 dataset as slope \* 0.01
(e.g., 12% slope = 0.12 PrecLossFrac). We acknowledge that this is a
simplification and that better parameterization would be possible if
soil types were available for each site. Tests showed that sites with
extreme slope tended to have lower biomass (Fig. 4), but the signal was
not that clear. The correlation between slope and elevation was 0.6
which could obscure the results.

Figure 4 Biomass of all species on each site sorted by slope, which was
used to modify the fraction of precipitation lost to run off (data from
replicate 1)

Some sites at very high elevation had, correspondingly, very low
temperature (Fig. 5). At these sites, the model showed a visible trend
in producing lower biomass (down to zero at a few sites see Fig. 6
below). Yet this was probably not apparent in the regeneration results,
as extremely low biomass could still support regeneration, even if
cohorts do not live long, due to our modification of sexual maturity
age.

Figure 5 Histogram of elevation of simulated sites.

Figure 6 Biomass of all species on each site ordered by elevation (data
from replicate 1)

@@

We have used the non-spatial seeding algorithm "universal dispersal" for
all the model runs. This way, every species can seed any forest site in
the landscape (across all the 200 sites). This mode does not take into
account the spatial arrangement and contiguity of cells (i.e., no active
seed dispersal), which is proper for the sake of the model comparison
exercise. Yet in real landscape seed limitation can also play a role in
species composition. The seed number is affected by species-specific age
of maturity, which we have modified (see Species section). This setting
ensures an influx of seeds from outside consistent with the original
settings. Simulation length We started each model run with initial
communities of all species at age 10 at all sites. Each run was
simulated for 1980 years. Sampling of model outputs was executed in two
time windows starting 50 years after the first cycle following typical
species longevity (see Species parameters). After the year 700, biomass
of most species was relatively stable and we sampled every 50 years,
starting at 750 until the year 1200. The next sampling period started
from 1450 and continued until 1900, i.e., after the next longevity peak
(Fig. 7), again every 50 years (see also Table 1). At each selected
sampling time, we extracted cohorts of each species according to the age
related to the two DBH thresholds (7cm/10cm) (see age-DBH relationships
in Table 2). Additionally, we extracted cohorts that were up to 9 years
older than these thresholds, to get the expected results for a 10-year
interval. For example, Fagus sylvatica reaches DBH 7cm at 17 years,
therefore we have selected all cohorts with age 17-26.

Figure 7 Biomass of each species in time across all sites (data from
replicate 1)

Figure 8 Sum of biomass of all species on each site in time (data from
replicate 1). Notice again the biomass decrease before approximately the
year 700 due to the effect of species longevity. The productivity of the
sites seem to be stable.

@@

We have sampled each site 20 times along the simulation run as described
in the simulation length chapter. To deliver more than one simulation
per site, we created 10 replicates with different climate data and
random number generator seed number to account for stochastic processes
in PnET-Succession. This resulted in the expected 200 samples for each
site. Output variables

sample Sample number code (1-200) was defined as sample in time +
(replicate-1)\*20, with sample in time (1-20) according to the sample
order. Replicate (1-10) combines random number and climate sequence (see
Table 1. For example sample number 25 was taken from the climate
replicate 2 and the fifth sample in time (year 950).

Table 1 Sample numbers and their corresponding settings

r.trees As LANDIS-II outputs are expressed in terms of cohort age and
mean biomass per site, we have calculated the mean DBH and biomass of
each species based on age. We estimated the age when each species
reaches 7 cm and 10 cm DBH and the number of individuals using
allometric models from the Italian forest inventory collected in R
package ForIT (Puletti et al. 2014), and unpublished data from the study
by Mina et al. (2021) (Table 2). Please note that the equations used
predict DBH \>0 already for a tree with age 1, this is because it is
based on forestry data and tree cores. The real age when reaching 1.3 m
height is species specific and we did not implement this offset. For
each cohort, we divided the cohort biomass reported by LANDIS-II,
recalculated to kg/ha by the mean biomass of a tree of the same species
and age to obtain the number of trees of each species per site. We
rounded all results up to the nearest integer to avoid any fraction of
trees. The number of individuals was aggregated for all the cohorts of
the age of reaching 7 cm or 10 cm DBH and all cohorts of the same
species that were up to 9 years older (Fig. 9). This approach does not
reflect influence of site conditions or stand densities on Age-DBH and
Age-Biomass. Relationships between age and DBH were derived by fitting
asymptotic regressions using individual tree measurements from an
ensemble of forest inventory plots distributed across south-eastern
Canada (see Supporting information of Mina et al. 2021). This was the
only public dataset for which we could relate tree age, derived from
stem coring and tree ring analysis, to measured DBH. Although tree age
measurements were obtained with a robust methodology, these functions do
not disentangle the influence of site conditions or stand densities.
Additionally, they reflect growing conditions of tree genera in Eastern
North America.

Figure 9 Number of trees in regeneration cohorts (both 7cm and 10 cm
groups) according to species. (data from all replicates and samples)

Table 2 Age at which each simulated species reaches DBH 7cm and 10 cm
Age at DBH 7cm Age at DBH 10cm Comment fasy (for Fagus sylvatica) 17 25\
piab (for Picea abies) 8 12\
abal (for Abies alba) 10 15\
cabe (for Carpinus betulus) 15 23 same equation as betu tico (for Tilia
cordata) 15 23 equation taken from quer acps (for Acer pseudoplatanus)
16 24\
betu (for Betula spp.) 15 23\
frex (for Fraxinus excelsior) 10 15\
quer (for Quercus spp.) 15 23\
algl (for Alnus glutinosa) 15 23 same equation as betu pisy (for Pinus
sylvestris) 9 13

r.ba basal area of regeneration Calculated as a sum of r.ba of all the
cohorts selected in r.trees using the following formula: DBH2 \* π/4 \*
r.tree (Fig. 10)..

Figure 10 Basal area of trees in regeneration cohorts (both 7cm and 10
cm groups) according to species. (data from all replicates and samples)

ba: total basal area We have not calculated this since our allometric
equations do not cover older ages. Therefore, we produced the following
two biomass variables as a surrogate.

@@

#### Runtime

Simulations on a 16 core virtual computer (each core 2.10 GHz) with 10
replicates running in parallel took about 3 hours 15 minutes. Processing
the results including allometry for selected cohorts took about an hour.
The model runs with a yearly timestep using monthly climate.

@@

Aber J. D. & Federer C. A. (1992). A generalized, lumped-parameter model
of photosynthesis, evapotranspiration and net primary production in
temperate and boreal forest ecosystems. Oecologia 92(4): 463-474. Aber
J.D., Ollinger S.V., Federer C.A., Reich P.B., Goulden M.L., Kicklighter
D.W., Melillo J.M. & Lathrop R.G. Jr. (1995) Predicting the effects of
climate change on water yield and forest production in the northeastern
United States. Climate Research 5(3): 207--22. de Bruijn A., Gustafson
E.J. , Sturtevant B.R., Foster J.R., Miranda B.R., Lichti N.I. & Jacobs
D.F. (2014) Toward more robust projections of forest landscape dynamics
under novel environmental conditions: Embedding PnET within LANDIS-II.
Ecological Modelling (287): 44--57 EEA (2019) Trends in atmospheric
concentrations of CO2 (ppm), CH4 (ppb) and N2O (ppb), between 1800 and
2017 available at
<https://www.eea.europa.eu/ds_resolveuid/217c026ca03649398aadf39d87623e57>
Gustafson, E. J., A. M. G. De Bruijn, R. E. Pangle, J.-M. Limousin, N.
G. McDowell, W. T. Pockman, B. R. Sturtevant, J. D. Muss, and M. E.
Kubiske. 2015. Integrating ecophysiology and forest landscape models to
improve projections of drought effects under climate change. Global
Change Biology 21:843-856. Mina, M., Messier, C., Duveneck, M., Fortin,
M. J., & Aquilué, N. 2021. Network analysis can guide resilience-based
management in forest landscapes under global change. Ecological
Applications, 31(1), 1--18. Petter G., Mairota P., Albrich K., Bebi P.,
Brůna J., Bugmann H., Haffenden A., Scheller R.M., Schmatz D.R., Seidl
R., Speich M., Vacchiano G. & Lischke H. (2020) How robust are future
projections of forest landscape dynamics? Insights from a systematic
comparison of four forest landscape models. Environmental Modelling &
Software 134: 104844. <https://doi.org/10.1016/j.envsoft.2020.104844>
Puletti N., Mura M., Castaldi C., Marchi N., Chiavetta U. and Scotti R.
(2014). ForIT: Functions from the 2nd Italian Forest Inventory (INFC). R
package version 1.0. <https://CRAN.R-project.org/package=ForIT>
Scheller, R. M., Domingo J. B. , Sturtevant B. R. , Williams J. S. ,
Rudy A. , Gustafson E. J. & Mladenoff D. J. (2007) Design, development,
and application of LANDIS-II, a spatial landscape simulation model with
flexible temporal and spatial resolution. Ecological Modelling 201(3-4):
409--419. <https://doi.org/10.1016/j.ecolmodel.2006.10.009> Tsubo M. &
Walker S. (2005) Relationships between photosynthetically active
radiation and clearness index at Bloemfontein, South Africa. Theoretical
and Applied Climatology 80: 17--25.
<https://doi.org/10.1007/s00704-004-0080-5>

>>>>

Landscape scale forest simulation model LANDIS-II (Scheller et al. 2007) with PnET-Succession extension (version 4.1). PnET-Succession is based on the Biomass Succession extension of Scheller and Mladenoff (2004), embedding elements of the PnET-II ecophysiology model of Aber et al. (1995) to mechanistically simulate growth as a competition for available light and water (De Bruijn et al. 2014). PnET-Succession simulates the competition among cohorts for water and light at a monthly time-step as a function of photosynthetic processes and maintenance respiration that are explicitly linked to environmental drivers such as temperature, precipitation, photosynthetic active radiation (PAR), and CO~2~ concentration. Regeneration depends on distance from a seed source, soil water, and subcanopy light, while competition is modeled by partitioning incoming light through multiple canopy layers. Growth of specific cohort biomass components is allocated in the following order: non structural carbon, first, then foliage then root and stem. Biomass growth decreases as cohorts approach their longevity age but mortality can occur at any time when carbon reserves production is insufficient to support growth due to shading, water competition, and/or drought (Gustafson et al. 2015), or by any disturbance or management extension used. For output, we selected Biomass Community Output Extension version 2.0.1, which generates output in the format: species age biomass (g / m^2^) at specified time steps.

@@

Climate input was generated by randomly selecting entire years from the monthly data from the CHELSA dataset and collating them one after the other to create a 2000 years series. The sampling was repeated 10 times with different random order of years to create the 10 climate replicates. Selecting a year at a time eliminates inconsistent seasons (especially summer). We used monthly mean, minimum and maximum temperatures and precipitation as input for PnET succession extension. PnET-Succession also requires mean monthly values of PAR during daylight hours, which we calculated by dividing the monthly rsds data by the number of daylight seconds in each month for latitude 47 to get W/m2. We then multiplied this by 4.6 to get μmol/m^2^/s and multiplied by 0.5 to get PAR part of the spectrum, based on the approximation by Tsubo et al. (2005). Leap years were not considered. The PAR distribution across all sites and years is in range with values that are typical for temperate regions.

CO~2~ concentration with yearly timestep was added from the concentrations table available from EEA (2019). For years 0-1750 CO~2~ concentration was maintained fixed to 278 ppm. As for the period 1750-1975 there was only one value each 5 years, we used a constant value for each 5-year period. CO~2~ concentration was added to the climate input file after randomization of temperature, precipitation and PAR (see above), so it reflects realistic values for the simulated years. Simulations were run from year 0 to 1990 but sampled outputs did not exceed year 1910. Thus, the drastic increase in CO~2~ concentration during the second half of the 1900s did not affect the delivered results.

@@

Soil type was set to LOAM at all sites. Rooting depth was calculated from soil quality codes (1-5, original values ranged from 1.13 to 4.87) multiplicated by 205. This produces a soil depth in the range 231 - 998 mm, that are within the typical range for PnET-Succession (maximum 1000 mm). A clear trend emerged - deeper soils resulted in higher biomass, although with some exceptions.

@@

The ecoregion-specific parameter Precipitation Loss Fraction denotes the water lost to runoff induced by topographic slope. We estimated this parameter for each site from topo_data_200 dataset as slope \* 0.01 (e.g., 12% slope = 0.12 PrecLossFrac). We acknowledge that this is a simplification and that better parameterization would be possible if soil types were available for each site. Tests showed that sites with extreme slope tended to have lower biomass, but the signal was not that clear. The correlation between slope and elevation was 0.6 which could obscure the results.

Some sites at very high elevation had, correspondingly, very low temperature. At these sites, the model showed a visible trend in producing lower biomass (down to zero at a few sites). Yet this was probably not apparent in the regeneration results, as extremely low biomass could still support regeneration, even if cohorts do not live long, due to our modification of sexual maturity age.

@@

As LANDIS-II cannot start from bare ground, initial communities were set to include all species defined in the protocol at all sites starting with 10-years old cohorts of the following species: fasy (*Fagus sylvatica*), pabi (*Picea abies*), abal (*Abies alba*), cabe (*Carpinus betulus*), tico (*Tilia cordata*), acps (*Acer pseudoplatanus*), betu (*Betula spp.*), frex (*Fraxinus excelsior*), quer (*Quercus spp.*), algl (*Alnus glutinosa*), pisy (*Pinus sylvestris*). The species-specific parameter denoting sexual maturity was set to 10 years equally for all species, so that seed source of all species is available from the onset of the simulations at all sites and to maintain seed in the ladscape pool. Vegetative regeneration (resprouting) was prevented, since this would have an effect on allometry and regeneration.

Some species may be under or overrepresented because their occurrence in LANDIS-II simulations is typically due to small-scale disturbances or management which were not included in the simulations.

The model requires two sets of parameters representing species-specific life traits. The first set is used by the LANDIS-II core module. The species-specific parameter denoting "longevity" and "sexual maturity" were the only parameters of this set considered in the current exercise. The "sexual maturity" was set to 10 years equally for all species, so that seed source of all species is available from the onset of the simulations at all sites and to maintain seed in the landscape pool. Vegetative regeneration (resprouting) was prevented since this would influence allometry and regeneration.

The second set is used by the PnET-Succession extension and includes several ecophysiological parameters. For this second set we started with the values already applied in Petter et al. (2020) with a previous version of PnET. The values of some of the parameters to which PnET is most sensitive and/or for which the range of variation had been modified for PnET4.1 were then iteratively adjusted (individually and then in combination) to meet the requirements of PnET version 4.1 used in this exercise. The changes were implemented one by one in simulations with the exercise data and those leading to an output deemed realistic and to a relative stabilization of the biomass trend through the simulation were retained for the implementation of the full set of replicates. The retained changes include: 1) decreased percentage of foliar nitrogen for algl (*Alnus glutinosa*); 2) rescaling to the new ranges for all species of the values of the Halfsat and the FracBelowG. The first of these parameters represents shade tolerance and its value affects the probability of establishment of a species (Half saturation light level for photosynthesis.); the latter regulates allocation between above and below ground pool (Fraction of non-foliar biomass that is belowground (root pool)) ; 3) a slight decrease of the value of the parameter (Fracfol) affecting the yearly allocation to foliage of the active total woody biomass for the three conifers abal (*Abies alba*), piab (*Picea abies*), pisy (*Pinus sylvestris*).

@@

Simulation area was defined as a rectangle of 10×20 dimension with sites (i.e., cells) of 1-ha. Each site is an individual ecoregion (i.e., one cell - as in our case - or group of cells sharing similar climate and soil conditions, as a way in LANDIS-II to define differences in biophysical conditions across landscapes) with specific values for soil quality, topography and climate, based on the data provided. The model runs with a yearly timestep using monthly climate. We started each model run with initial communities of all species at age 10 at all sites. Each run was simulated for 1980 years.

@@

We have used the non-spatial seeding algorithm "universal dispersal" for all the model runs. This way, every species can seed any forest site in the landscape (across all the 200 sites). This mode does not take into account the spatial arrangement and contiguity of cells (i.e., no active seed dispersal), which is proper for the sake of the model comparison exercise. Yet in real landscape seed limitation can also play a role in species composition. The seed number is affected by species-specific age of maturity, which we have modified. This setting ensures an influx of seeds from outside consistent with the original settings.

@@

We have sampled each site 20 times along the simulation run in two time windows starting 50 years after the first cycle following typical species longevity. After the year 700, biomass of most species was relatively stable and we sampled every 50 years, starting at 750 until the year 1200. The next sampling period started from 1450 and continued until 1900, i.e., after the next longevity peak, again every 50 years. At each selected sampling time, we extracted cohorts of each species according to the age related to the two DBH thresholds (7cm/10cm). Additionally, we extracted cohorts that were up to 9 years older than these thresholds, to get the expected results for a 10-year interval. For example, *Fagus sylvatica* reaches DBH 7cm at 17 years, therefore we have selected all cohorts with age 17-26.

To deliver more than one simulation per site, we created 10 replicates with different climate replicate and random number generator seed number to account for stochastic processes in PnET-Succession. This resulted in the expected 200 samples for each site.

*Please describe how have you calculated each of the variables in the output data.*

Sample number code

Number (1-200) was defined as sample in time + (replicate-1)\*20, with sample in time (1-20) according to the sample order. Replicate (1-10) combines random number and climate sequence. For example sample number 25 was taken from the climate replicate 2 and the fifth sample in time (year 950).

r.trees

As LANDIS-II outputs are expressed in terms of cohort age and mean biomass per site, we have calculated the mean DBH and biomass of each species based on age. We estimated the age when each species reaches 7 cm and 10 cm DBH and the number of individuals using allometric models from the Italian forest inventory collected in R package ForIT (Puletti et al. 2014), and unpublished data from the study by Mina et al. (2021). Please note that the equations used predict DBH \>0 already for a tree with age 1, this is because it is based on forestry data and tree cores. The real age when reaching 1.3 m height is species specific and we did not implement this offset. For each cohort, we divided the cohort biomass reported by LANDIS-II, recalculated to kg/ha by the mean biomass of a tree of the same species and age to obtain the number of trees of each species per site. We rounded all results up to the nearest integer to avoid any fraction of trees. The number of individuals was aggregated for all the cohorts of the age of reaching 7 cm or 10 cm DBH and all cohorts of the same species that were up to 9 years older. This approach does not reflect influence of site conditions or stand densities on Age-DBH and Age-Biomass. Relationships between age and DBH were derived by fitting asymptotic regressions using individual tree measurements from an ensemble of forest inventory plots distributed across south-eastern Canada (see Supporting information of Mina et al. 2021). This was the only public dataset for which we could relate tree age, derived from stem coring and tree ring analysis, to measured DBH. Although tree age measurements were obtained with a robust methodology, these functions do not disentangle the influence of site conditions or stand densities. Additionally, they reflect growing conditions of tree genera in Eastern North America.

r.ba basal area of regeneration Calculated as a sum of r.ba of all the cohorts selected in r.trees using the following formula: DBH2 \* π/4 \* r.tree.

ba: total basal area We have not calculated this since our allometric equations do not cover older ages. Therefore, we produced the following two biomass variables as a surrogate. The basal area was later calculated as a sum of r.ba and estimated ba based on biomass. The estimation includes transformation of kg into tonnes / ha and expansion factor 12.5 to transform ba m^2^/ha.

((( landis\$sum.bio - landis\$r.bio.sum )\* 0.001) / 12.5). This includes

@@

@@

Aber J. D. & Federer C. A. (1992). A generalized, lumped-parameter model of photosynthesis, evapotranspiration and net primary production in temperate and boreal forest ecosystems. Oecologia 92(4): 463-474.

Aber J.D., Ollinger S.V., Federer C.A., Reich P.B., Goulden M.L., Kicklighter D.W., Melillo J.M. & Lathrop R.G. Jr. (1995) Predicting the effects of climate change on water yield and forest production in the northeastern United States. Climate Research 5(3): 207--22.

de Bruijn A., Gustafson E.J. , Sturtevant B.R., Foster J.R., Miranda B.R., Lichti N.I. & Jacobs D.F. (2014) Toward more robust projections of forest landscape dynamics under novel environmental conditions: Embedding PnET within LANDIS-II. Ecological Modelling (287): 44--57

EEA (2019) Trends in atmospheric concentrations of CO2 (ppm), CH4 (ppb) and N2O (ppb), between 1800 and 2017 available at <https://www.eea.europa.eu/ds_resolveuid/217c026ca03649398aadf39d87623e57>

Gustafson, E. J., A. M. G. De Bruijn, R. E. Pangle, J.-M. Limousin, N. G. McDowell, W. T. Pockman, B. R. Sturtevant, J. D. Muss, and M. E. Kubiske. 2015. Integrating ecophysiology and forest landscape models to improve projections of drought effects under climate change. Global Change Biology 21:843-856.

Mina, M., Messier, C., Duveneck, M., Fortin, M. J., & Aquilué, N. 2021. Network analysis can guide resilience-based management in forest landscapes under global change. Ecological Applications, 31(1), 1--18.

Petter G., Mairota P., Albrich K., Bebi P., Brůna J., Bugmann H., Haffenden A., Scheller R.M., Schmatz D.R., Seidl R., Speich M., Vacchiano G. & Lischke H. (2020) How robust are future projections of forest landscape dynamics? Insights from a systematic comparison of four forest landscape models. Environmental Modelling & Software 134: 104844. <https://doi.org/10.1016/j.envsoft.2020.104844>

Puletti N., Mura M., Castaldi C., Marchi N., Chiavetta U. and Scotti R. (2014). ForIT: Functions from the 2nd Italian Forest Inventory (INFC). R package version 1.0. <https://CRAN.R-project.org/package=ForIT>

Scheller, R. M., Domingo J. B. , Sturtevant B. R. , Williams J. S. , Rudy A. , Gustafson E. J. & Mladenoff D. J. (2007) Design, development, and application of LANDIS-II, a spatial landscape simulation model with flexible temporal and spatial resolution. Ecological Modelling 201(3-4): 409--419. <https://doi.org/10.1016/j.ecolmodel.2006.10.009>

Tsubo M. & Walker S. (2005) Relationships between photosynthetically active radiation and clearness index at Bloemfontein, South Africa. Theoretical and Applied Climatology 80: 17--25. <https://doi.org/10.1007/s00704-004-0080-5>
